---
title: "Untitled"
author: "PRU"
date: "16/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries
library(tidyverse) # catch all
```

read in all data

```{r}
# just a list of files inside master.zip
zip_archive <- "../__DATA/BRA/BRA_2019.zip"
archive     <- as.character(unzip(zip_archive, list = TRUE)$Name)

# loop over all files and read them in
ds <- purrr::map(
   .x = archive
  ,.f = ~ read_csv( unz(zip_archive, .x), col_types = cols(.default = "c"))
  )

ds <- ds %>% map(
  .f = ~ mutate(
    .x
    ,PHASE = case_when(
         length(unique(ADEP)) < length((unique(ADES))) ~ "DEP" #ADEP[1]
        ,length(unique(ADES)) < length((unique(ADEP))) ~ "ARR" # ADES[1]
        ,TRUE ~ NA_character_)
    )
) %>% bind_rows
```

split arrivals and departure

```{r}
arrs <- ds %>% filter(PHASE == "DEP")
deps <- ds %>% filter(PHASE == "ARR")
```

identify busiest (departure) airport pair

```{r}
cps <- deps %>% 
  group_by(ADEP, ADES) %>% 
  summarise(N = n(), .groups = "drop")

# let's pick the top 3 for some airports of the study
cps_top <- cps %>% 
  filter(ADEP %in% c("SBSP","SBRJ","SBBR","SBGR") ) %>% 
  group_by(ADEP) %>%  
  slice_max(N, n = 3)

cps_top
```

top-pair: SBRJ	SBSP	18937

```{r}
adep <- "SBRJ"
ades <- "SBSP"

pair_airport_flights <- function(.adep, .ades, .ds = ds, .threshold = 180){
  # break out departures and arrivals ------------------------------------------
  deps <- .ds %>% filter(ADEP == .adep, PHASE == "DEP") %>% 
    mutate(DOF = lubridate::date(AOBT)) %>% 
    select(ADEP, ADES, FLTID, CLASS, TYPE, DRWY, AOBT, ATOT)
  
  arrs <- .ds %>% filter(ADES == .ades, PHASE == "ARR") %>% 
    mutate(DOF = lubridate::date(AIBT)) %>% 
    select(ADEP, ADES, FLTID, CLASS, TYPE, ARWY = DRWY, ALDT, AIBT)
  
  # bind departures and arrivals -----------------------------------------------
  flts <- deps %>% 
    left_join(arrs, by = c("ADEP", "ADES", "FLTID", "CLASS", "TYPE")) %>% 
    # coerce timestamps
    mutate(across( .cols = c("AOBT","ATOT","ALDT","AIBT")
                  , .fns = lubridate::ymd_hms )
           ,AIR_TIME = difftime(ALDT, ATOT, units = "min")
           ,FLT_TIME = difftime(AIBT, AOBT, units = "min")
           )
  
  # filter erronuous combinations 
  # a flight must have positive flight duration and an expected duration < threshold
  flts <- flts %>% 
    filter(AIR_TIME > 0, AIR_TIME <= .threshold)
}
```

```{r}
rq <- pair_airport_flights(adep, ades)

rq %>% filter(AIR_TIME > 0, AIR_TIME <= 360) %>% 
  ggplot() + 
    geom_histogram(aes(x = AIR_TIME), binwidth = 5) + 
    scale_x_continuous(labels = scales::comma) +
  theme_minimal()
```

let's check for the most frequent connections

```{r}
freq_flts <- rq %>% 
  group_by(FLTID) %>% 
  summarise(N = n()
            , MAX = max(AIR_TIME)
            , MIN = min(AIR_TIME)
            , MEAN= mean(AIR_TIME)
            , .groups = "drop") %>% 
  arrange(desc(N))

freq_flts
```

getting a feel for the sample

```{r}
# sig. number of connections, e.g. 5 per month ~ 60, 10 ~ 120, 20 ~ 240
cons <- c(60, 120, 240)

freq_flts %>% 
  ggplot() + 
    geom_step(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = N, group = "all")) + 
    geom_hline(yintercept = cons, linetype = "dotdash") +
    theme_minimal()
```

```{r}
tmp <- freq_flts %>% arrange(desc(N)) %>% mutate(RANK = row_number())

rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = AIR_TIME, colour = GROUP)) +
    theme_minimal() +
  labs(title = "Air-time (ATOT-ALDT) distribution - all flights")
```

```{r}
rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = FLT_TIME, colour = GROUP)) +
    theme_minimal() +
  labs(title = "Block-time (AOBT - AIBT) distribution - all flights")
```

```{r}
rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  filter(GROUP %in% c("(60,120]","(120,240]","(240,Inf]")) %>% 
  
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = FLT_TIME, colour = GROUP)) +
    theme_minimal()
```

```{r}
flight_var <- rq %>% 
  mutate( TXOT = difftime(ATOT, AOBT, units = "min")
         ,TXIT = difftime(ATOT, AOBT, units = "min")
         ) %>% 
  left_join(tmp %>% select(FLTID, N, RANK), by = "FLTID") 
```

read in all data

#```{r}
# just a list of files inside master.zip
zip_archive <- "../__DATA/BRA/BRA_2019.zip"
archive     <- as.character(unzip(zip_archive, list = TRUE)$Name)

 #loop over all files and read them in

ds <- purrr::map(
  .x = archive
  ,.f = ~ read_csv( unz(zip_archive, .x), col_types = cols(.default = "c"))
  )

ds <- ds %>% map(.f = ~ mutate(
  .x
    ,PHASE = case_when(
         length(unique(ADEP)) < length((unique(ADES))) ~ "DEP" #ADEP[1]
        ,length(unique(ADES)) < length((unique(ADEP))) ~ "ARR" # ADES[1]
        ,TRUE ~ NA_character_)
    )
%>% bind_rows
```

split arrivals and departure

#```{r}
arrs <- ds %>% filter(PHASE == "DEP")
deps <- ds %>% filter(PHASE == "ARR")
```

identify busiest (departure) airport pair

#```{r}
cps <- deps %>% 
  group_by(ADEP, ADES) %>% 
  summarise(N = n(), .groups = "drop")

# let's pick the top 3 for some airports of the study
cps_top <- cps %>% 
  filter(ADEP %in% c("SBSP","SBRJ","SBBR","SBGR") ) %>% 
  group_by(ADEP) %>%  
  slice_max(N, n = 3)

cps_top
```

top-pair: SBRJ	SBSP	18937

#```{r}
adep <- "SBRJ"
ades <- "SBSP"

pair_airport_flights <- function(.adep, .ades, .ds = ds, .threshold = 180){
  # break out departures and arrivals ------------------------------------------
  deps <- .ds %>% filter(ADEP == .adep, PHASE == "DEP") %>% 
    mutate(DOF = lubridate::date(AOBT)) %>% 
    select(ADEP, ADES, FLTID, CLASS, TYPE, DRWY, AOBT, ATOT)
  
  arrs <- .ds %>% filter(ADES == .ades, PHASE == "ARR") %>% 
    mutate(DOF = lubridate::date(AIBT)) %>% 
    select(ADEP, ADES, FLTID, CLASS, TYPE, ARWY = DRWY, ALDT, AIBT)
  
  # bind departures and arrivals -----------------------------------------------
  flts <- deps %>% 
    left_join(arrs, by = c("ADEP", "ADES", "FLTID", "CLASS", "TYPE")) %>% 
    # coerce timestamps
    mutate(across( .cols = c("AOBT","ATOT","ALDT","AIBT")
                  , .fns = lubridate::ymd_hms )
           ,TXOT     = difftime(ATOT, AOBT, units = "min")
           ,AIR_TIME = difftime(ALDT, ATOT, units = "min")
           ,TXIT     = difftime(AIBT, ALDT, units = "min")
           ,FLT_TIME = difftime(AIBT, AOBT, units = "min")
           )
  
  # filter erronuous combinations 
  # a flight must have positive flight duration and an expected duration < threshold
  flts <- flts %>% 
    filter(AIR_TIME > 0, AIR_TIME <= .threshold)
}
```

#```{r}
rq <- pair_airport_flights(adep, ades) %>% 
  mutate(YEAR = lubridate::year(ATOT))

rq %>% filter(AIR_TIME > 0, AIR_TIME <= 360) %>% 
  ggplot() + 
    geom_histogram(aes(x = AIR_TIME), binwidth = 5) + 
    scale_x_continuous(labels = scales::comma) +
  theme_minimal()
```

let's check for the most frequent connections

#```{r}
freq_flts <- rq %>% 
  group_by(FLTID) %>% 
  summarise(N = n()
            , MAX = max(AIR_TIME)
            , MIN = min(AIR_TIME)
            , MEAN= mean(AIR_TIME)
            , .groups = "drop") %>% 
  arrange(desc(N))

freq_flts
```

getting a feel for the sample

#```{r}
# sig. number of connections, e.g. 5 per month ~ 60, 10 ~ 120, 20 ~ 240
cons <- c(60, 120, 240)

freq_flts %>% 
  ggplot() + 
    geom_step(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = N, group = "all")) + 
    geom_hline(yintercept = cons, linetype = "dotdash") +
    theme_minimal()
```

#```{r}
tmp <- freq_flts %>% arrange(desc(N)) %>% mutate(RANK = row_number())

rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = AIR_TIME, colour = GROUP)) +
    theme_minimal() +
  labs(title = "Air-time (ATOT-ALDT) distribution - all flights")
```

#```{r}
rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = FLT_TIME, colour = GROUP)) +
    theme_minimal() +
  labs(title = "Block-time (AOBT - AIBT) distribution - all flights")
```

#```{r}
rq %>% 
  left_join(tmp %>% select(FLTID, N, RANK)) %>% 
  mutate(GROUP = cut(N, breaks = c(0,60,120,240, Inf))) %>% 
  filter(GROUP %in% c("(60,120]","(120,240]","(240,Inf]")) %>% 
  
  ggplot() + 
    geom_boxplot(aes(x = forcats::fct_reorder(FLTID, desc(N)), y = FLT_TIME, colour = GROUP)) +
    theme_minimal()
```

#```{r}
calc_flight_var <- function(.flts, .threshold = 240){
 ranking <- .flts %>% 
   group_by(YEAR, ADEP, ADES, FLTID) %>% 
   summarise(N = n(), .groups = "drop") %>% 
   group_by(ADEP) %>% 
   arrange(desc(N), .by_group = TRUE) %>% 
   mutate(RANK = row_number())
  
  flight_var <- .flts %>% 
    left_join(ranking, by = c("YEAR", "ADEP", "ADES" ,"FLTID")) %>% 

    filter(N >= .threshold) %>% 
    group_by(ADEP, ADES, YEAR, FLTID) %>% 
    summarise( N = n()
              ,PCT_15 = quantile(FLT_TIME, probs = 0.15, na.rm = TRUE) %>% as.numeric()
              ,PCT_85 = quantile(FLT_TIME, probs = 0.85, na.rm = TRUE) %>% as.numeric()
              ,PCT_20 = quantile(FLT_TIME, probs = 0.20, na.rm = TRUE) %>% as.numeric()
              ,PCT_80 = quantile(FLT_TIME, probs = 0.80, na.rm = TRUE) %>% as.numeric()
              ,.groups = "drop")
}
```

#```{r}
flt_var <- calc_flight_var(rq)
flt_var
```

# Prepare Analytical Data

#```{r}
# just a list of files inside master.zip ======================================
yr <- 2021
zip_archive <- paste0("../__DATA/BRA/BRA_", yr, ".zip")
archive     <- as.character(unzip(zip_archive, list = TRUE)$Name)

# loop over all files and read them in
ds <- purrr::map(
   .x = archive
  ,.f = ~ read_csv( unz(zip_archive, .x), col_types = cols(.default = "c"))
  )

ds <- ds %>% map(
  .f = ~ mutate(
    .x
    ,PHASE = case_when(
         length(unique(ADEP)) < length((unique(ADES))) ~ "DEP" #ADEP[1]
        ,length(unique(ADES)) < length((unique(ADEP))) ~ "ARR" # ADES[1]
        ,TRUE ~ NA_character_)
    )
) %>% bind_rows
#---------------------------------- end load data -----------------------------

# ds keeps all study data for Region
arrs <- ds %>% filter(PHASE == "DEP")
deps <- ds %>% filter(PHASE == "ARR")

# study airports
bra_apts <- c("SBBR", "SBGR", "SBSP", "SBKP", "SBRJ", "SBGL", "SBCF", "SBSV", "SBPA","SBCT")
arrs <- arrs %>% filter(ADES %in% bra_apts)
deps <- deps %>% filter(ADEP %in% bra_apts)

cps <- deps %>% 
  group_by(ADEP, ADES) %>% 
  summarise(N = n(), .groups = "drop") %>% 
  group_by(ADEP) %>% 
  slice_max(N, n = 10) # identify top n city pairs

pairs <- cps %>% 
  select(-N) %>% 
  purrr::pmap_dfr(.f = ~ pair_airport_flights(..1, ..2))
```

write_csv(pairs, paste0("./data/BRA_cps_", yr, ".csv"))

# Performance Analytics

#```{r}
fns <- list.files(path = "./data/", pattern = "BRA_cps", full.names = TRUE)

pairs <- fns %>% 
  purrr::map_dfr(.f = ~ read_csv(.x)) %>% 
  mutate(YEAR = lubridate::year(ATOT))

kpi_15_comp <- calc_flight_var(pairs, .threshold = 120) %>% 
  filter(ADEP != ADES)     # eleminate local flights

kpi_15 <- kpi_15_comp %>% 
  mutate( VAR_1585 = (PCT_85 - PCT_15) / 2 
         ,VAR_2080 = (PCT_80 - PCT_20) / 2 
         ,N_VAR_1585 = N * VAR_1585
         ,N_VAR_2080 = N * VAR_2080
         ) %>% 
  group_by(YEAR, ADEP, ADES) %>% 
  summarise( FLTS = sum(N)
            ,FLT_VAR_1585 = sum(N_VAR_1585) / sum(N)
            ,FLT_VAR_2080 = sum(N_VAR_2080) / sum(N)
            ,.groups = "drop")
```

``{r, warning=FALSE, message=FALSE}
`{r, warning=FALSE, message=FALSE}
rning=FALSE, message=FALSE}
kpi_15 %>% 
  ggplot() +
    geom_col (aes(x = YEAR, y = FLT_VAR_1585, fill  = ADES, colour = ADES)) +
    geom_step(aes(x = YEAR, y = FLT_VAR_1585, group = ADES, colour = ADES)) +
    facet_grid(ADEP ~ ADES) + 
  theme_minimal()
```

https://stackoverflow.com/questions/60104140/splitting-a-facet-wrap-into-two-additional-facets


#```{r}
add_ym <- function(.df, .timestamp = ATOT){
  df_ym <- .df %>% 
    mutate( DOF = lubridate::date({{.timestamp}})
           ,YM  = lubridate::format_ISO8601(DOF, precision = "ym")
           )
}

var_timeline <- pairs %>% 
  filter(ADEP == "SBBR", ADES == "SBSP") %>% 
  #add_ym() %>%
  mutate(BIN = lubridate::floor_date(ATOT, unit = "month")) %>% 
  group_by(ADEP, ADES, BIN) %>% 
  summarise( N = n()
            ,AVG_TXOT = median(TXOT, na.rm = TRUE)
            ,AVG_AIR  = median(AIR_TIME, na.rm = TRUE)
            ,AVG_TXIT = median(TXIT, na.rm = TRUE)
            ,.groups = "drop")

var_timeline
```

#```{r}
var_timeline %>% 
  tidyr::pivot_longer(cols = contains("AVG_"), names_to = "PHASE", values_to = "TIME") %>% 
  ggplot() +
    geom_line(aes(x = BIN, y = TIME, group = PHASE, linetype = PHASE, colour = PHASE))
```

#```{r}
var_timeline_add <- pairs %>% 
  filter(ADEP == "SBBR", ADES == "SBSP") %>% 
  mutate( REF_TXOT = quantile(TXOT, probs = 0.2, na.rm = TRUE), ADD_TXOT = TXOT - REF_TXOT
         ,REF_AIR  = quantile(AIR_TIME, probs = 0.2, na.rm = TRUE), ADD_AIR = AIR_TIME - REF_AIR
         ,REF_TXIT = quantile(TXIT, probs = 0.2, na.rm = TRUE), ADD_TXIT = TXIT - REF_TXIT
         ,BIN      = lubridate::floor_date(ATOT, unit = "month")
         ) %>% 
  group_by(ADEP, ADES, BIN) %>% 
  summarise( N = n()
            ,ADD_TXOT = median(ADD_TXOT, na.rm = TRUE)
            ,ADD_AIR  = median(ADD_AIR, na.rm = TRUE)
            ,ADD_TXIT = median(ADD_TXIT, na.rm = TRUE)
            ,.groups = "drop")

var_timeline_add

```

#```{r}
var_timeline_add %>% 
  tidyr::pivot_longer(cols = contains("ADD_"), names_to = "PHASE", values_to = "TIME") %>% 
  ggplot() +
    geom_line(aes(x = BIN, y = TIME, group = PHASE, linetype = PHASE, colour = PHASE))
```

#```{r}
monthly_var_by_phase <- pairs %>% 
    filter(ADEP == "SBBR", ADES == "SBSP") %>% 
    mutate(MOF = lubridate::floor_date(ATOT, unit = "month")) %>% 
    group_by(ADEP, ADES, MOF) %>% 
    summarise( N = n()
              ,TXOT_P15 = quantile(TXOT, probs = 0.15, na.rm = TRUE)
              ,TXOT_P85 = quantile(TXOT, probs = 0.85, na.rm = TRUE)
              ,AIR_P15 = quantile(AIR_TIME, probs = 0.15, na.rm = TRUE)
              ,AIR_P85 = quantile(AIR_TIME, probs = 0.85, na.rm = TRUE)
              ,TXIT_P15 = quantile(TXIT, probs = 0.15, na.rm = TRUE)
              ,TXIT_P85 = quantile(TXIT, probs = 0.85, na.rm = TRUE)
              ,.groups = "drop") %>% 
    mutate( VAR_TXOT = (TXOT_P85 - TXOT_P15) / 2
           ,VAR_AIR  = (AIR_P85  - AIR_P15)  / 2
           ,VAR_TXIT = (TXIT_P85 - TXIT_P15) / 2
           )

monthly_var_by_phase
```

#```{r}
p1 <- monthly_var_by_phase %>% 
  tidyr::pivot_longer(
    cols = starts_with("VAR_"), names_to = "PHASE", values_to = "VAR") %>% 
  mutate(YEAR = lubridate::year(MOF)
        , PHASE = factor(PHASE, levels = c("VAR_TXOT", "VAR_AIR", "VAR_TXIT"))) %>% 
  

  ggplot() + 
  geom_line(aes(x = MOF, y = VAR, group = PHASE, colour = PHASE )) +
  scale_colour_discrete( name= "phase"
                     ,breaks = c("VAR_AIR", "VAR_TXOT", "VAR_TXIT")
                     ,labels = c("airborne", "taxi-out", "taxi-in")
                     ) +
  theme_minimal() +
  theme(legend.position = c(0.9, 0.8)) +
  labs(x = NULL, y = "variability (min)", colour = "phase")
p1
```

#```{r}
p2 <- monthly_var_by_phase %>% 
  ggplot() + geom_line(aes(x = MOF, y = N)) +
  theme_minimal() +
  labs(x = NULL, y = "flights")

library(patchwork)
(p2 / p1) + 
  plot_layout(heights = c(1,3)) +
  plot_annotation(title = "SBBR-SBSP - Flight Time Variability per Flight Phase")
```

#```{r}
p1 + facet_grid(. ~ PHASE) +
  coord_cartesian(clip = "off") +
theme(panel.spacing.x = unit(6, "mm"))
```

# Network level analysis

#```{r}
cps_top <- pairs %>% 
  group_by(ADEP, ADES, YEAR) %>% 
  summarise(N = n(), .groups = "drop") %>% 
  filter(ADEP != ADES)   # remove round robins

cps_top_2019 <- cps_top %>% 
  filter(YEAR == 2019) %>% 
  group_by(ADEP) %>% 
  slice_max(N, n = 5)

tops <- pairs %>% inner_join(cps_top_2019 %>% select(ADEP, ADES)) 

monthly_var_by_phase <- 
  
  tops %>% 
  mutate(MOF = lubridate::floor_date(ATOT, unit = "month")) %>% 
    group_by(ADEP, ADES, MOF) %>% 
    summarise( N = n()
              ,TXOT_P15 = quantile(TXOT, probs = 0.15, na.rm = TRUE)
              ,TXOT_P85 = quantile(TXOT, probs = 0.85, na.rm = TRUE)
              ,AIR_P15 = quantile(AIR_TIME, probs = 0.15, na.rm = TRUE)
              ,AIR_P85 = quantile(AIR_TIME, probs = 0.85, na.rm = TRUE)
              ,TXIT_P15 = quantile(TXIT, probs = 0.15, na.rm = TRUE)
              ,TXIT_P85 = quantile(TXIT, probs = 0.85, na.rm = TRUE)
              ,.groups = "drop") %>% 
    mutate( VAR_TXOT = (TXOT_P85 - TXOT_P15) / 2
           ,VAR_AIR  = (AIR_P85  - AIR_P15)  / 2
           ,VAR_TXIT = (TXIT_P85 - TXIT_P15) / 2
           )
```

#```{r}
tmp <- monthly_var_by_phase %>% 
  tidyr::pivot_longer(
    cols = starts_with("VAR_"), names_to = "PHASE", values_to = "VAR") %>% 
  mutate(YEAR = lubridate::year(MOF)
        , PHASE = factor(PHASE, levels = c("VAR_TXOT", "VAR_AIR", "VAR_TXIT"), labels = c("taxi-out", "airborne", "taxi-in"))
        , CP = paste0(ADEP, "-", ADES)) 
    #  %>% filter(CP == "SBBR-SBRJ")  

tmp_med <- tmp %>% group_by(MOF, PHASE) %>% 
  summarise(VAR = median(VAR, na.rm = TRUE)
            ,VAR_MN = mean(VAR, na.rm = TRUE))

p1 <-monthly_var_by_phase %>% 
  group_by(MOF) %>% summarise(N = sum(N)) %>% 
  ggplot() +
  geom_line(aes(x = MOF, y = N)) +
  theme_minimal()

p2 <- 
    ggplot() +
    geom_line(data = tmp, aes(x = MOF, y = VAR, group = CP), colour = "grey80") + 
    geom_line(data = tmp_med,aes(x = MOF, y = VAR, group = "median"), size = 2 , colour = "blue") +
  geom_line(data = tmp_med,aes(x = MOF, y = VAR_MN, group = "mean"), size = 1, linetype = "dotdash" , colour = "green") +
    facet_grid(. ~ PHASE) +
  theme_minimal() +
  theme( legend.position = c(0.9, 0.8)
        ,panel.spacing.x = unit(5, "mm")
        ,plot.margin = margin(0, 5, 0, 0, unit = "mm")) +
  labs(x = NULL, y = "time variability [min]"
       ,title = "Regional Flight Time Variability - Brazil"
       ,caption = "Top 5 connections per study airport.")

#p1 / p2

p2

```


FUEL BURN
Note: thismight not be totally in sync with what we did above.

* loaded city pairs
* summarised it



