---
title: "Brazil-Europe"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#FFFFFF"   # white
      fg: "#111111" 
      primary: "#2489E7"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

<!-- 
THIS IS A RMarkdown document in a quarto project!
To render the dashboard use the console!

rmarkdown::render("./dashboard/bra-eur-db.Rmd")

-->

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

# Landing Page

Column {data-width=650 .tabset}
-----------------------------------------------------------------------

### Chart A

```{r}
fn <- here::here("figures","scope-airports.png")
knitr::include_graphics(fn)
```

The bi-regional comparison assesses performance at a subset of the airports.

### Chart B

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart C

```{r}

```

### Chart D

```{r}

```

# Next Page

### Articles per Day

```{r}
articles <- 25
valueBox(articles, icon = "fa-pencil")
```

### Comments per Day

```{r}
comments <- 45
valueBox(comments, icon = "fa-comments")
```

### Spam per Day

```{r}
spam <- 78
valueBox(spam, 
         icon = "fa-trash",
         color = ifelse(spam > 10, "warning", "primary"))
```

# About

The Brazil-Europe Operational ANS Performance Comparison is a joint project by the Performance Section of the Brazilian Airspace Control Department (DECEA) and the Performance Review Unit of EUROCONTROL.

The project aims to foster the understanding of operational performance in both regions on the basis of a set of harmonised performance indicators, supporting methods, and data.